Hyperplane Elimination for Quickly Enumerating Local Optima

### 2015-08-26 Hyperplane Enumeration/main.tex
Added a bunch more, fixed line ending problems

% first the title is needed
\title{Hyperplane Elimination for Quickly Enumerating MK Landscapes}

@@

%The abstract should summarize the contents of the paper and should
%contain at least 70 and at most 150 words. It should be written using the
%\emph{abstract} environment.
\keywords{Landscape Understanding, Gray-Box}
\end{abstract}

\section{Introduction}
TODO Motivate using network stuff.

\section{MK Landscapes and Gray-Box Optimization}
TODO Explain all problems here

@@

\section{Gray-Box Enumeration}
TODO Explain how gray-box properties and gray coding can be used to
enumerate in \BigO{2^N} time.

@@

\section{Timing Comparisons}

>>>>

% first the title is needed
\title{Hyperplane Elimination for Quickly Enumerating Local Optima}

@@

%The abstract should summarize the contents of the paper and should
%contain at least 70 and at most 150 words. It should be written using the
%\emph{abstract} environment.
\keywords{Landscape Understanding, Gray-Box, MK Landscapes}
\end{abstract}

\section{Introduction}
The ruggedness and high dimensionality of most interesting landscapes makes them challenging
to visualize or otherwise analyze. However, doing so can be helpful in
quantifying the difficulty of a problem, and how to best design algorithms to deal
with those difficulties. 
Similarly, knowing which characteristics favor
a particular
algorithm can help researchers choose the algorithm
most likely to perform well on their problem.

One avenue for performing this problem understanding is to
examine the frequency and distribution of local optima~\cite{boese:1994:bigvalley},
and how operators allow search to transition between these
optima~\cite{tomassini:2008:nknetworks,verel:2011:nknetworks,ochoa:2015:crossovernetworks}.
However, finding all of the local optima is prohibitively time consuming
for even small problems, with many studies limited to 18-bit
problems~\cite{tomassini:2008:nknetworks,verel:2011:nknetworks}.
By leveraging recent advancements in Gray-Box optimization which allow for
constant time local search~\cite{chicano:2014:ball}, this limit was raised
to 30-bit problems~\cite{ochoa:2015:crossovernetworks}.

Here we will introduce a method for finding all local optima of problems
with up to 80 bits. The cornerstone of this method is the identification
of hyperplanes which cannot contain any local optima, allowing large
portions of the search space to be skipped during enumeration.

\section{Mk Landscapes and Gray-Box Optimization}
An Mk Landscape is any function which can be expressed in the following form:
\begin{equation}
  f(x) = \sum_{i=1}^{M} f_i(mask(x, s_i))
  \label{eq-mk}
\end{equation}
In this equation there are $M=$\BigO{N} subfunctions $f_i$, which act
on some subset $s_i$ of variables in $x$. Each subfunction performs the mapping
$f_i : \mathbb{B}^{k_i}\rightarrow \mathbb{R}$ where $k_i \leq k$ and $k$ is
a constant with respect to $N$. The $mask$ function returns the values of $x$ for
the variables in $s_i$.

A critical aspect of this problem class is that it can be optimized using Gray-Box
techniques, which can allow for highly efficient local~\cite{whitley:2013:greedy}
and global~\cite{goldman:2015:GBO,tintos:2015:partitioncross} search. These algorithms
exploit the known and limited non-linearity between variables expressed by all $s_i$,
as well as perform partial solution evaluation using all $f_i$.

This formulation can represent many problems of real-world interest, as well as
the most commonly used combinatorial benchmark problems. In this work we shall
examine 5 problems in particular: Concatenated Traps, Adjacent and Random NKq Landscapes,
Ising Spin Glasses, and MAX-kSAT.

The Concatenated Trap problem~\cite{deb:1992:trap} is a composition of $k$-order deceptive
separable subfunctions. In Mk Landscape terms, $M=N/k$ such that $\forall_i |s_i| = k$ and
$\forall_{i \neq j} s_i \cap s_j = \emptyset$. Each $f_i$ applies an identical function based
on the number of variables set to 1:
\begin{equation}
   trap(t) = \left\{
     \begin{array}{rl}
       k-1-t, &  t<k\\
       k,   &  t = k
     \end{array}
   \right.
  \label{eq-trap}
\end{equation}
While this problem is still in common use~\cite{hsu:2015:dsmgaII,inoue:2015:adaptivep3},
most advanced methods can solve it trivially~\cite{goldman:2012:ltga} and when expressed
as an Mk Landscape can be solved exactly in \BigO{N} time~\cite{whitley:2015:mk}. However,
we include it here as its simple structure allows for straight forward algorithm analysis.

TODO Explain all problems here

@@

\section{Gray-Box Enumeration}
TODO Explain how gray-box properties and gray coding can be used to
enumerate in \BigO{2^N} time.~\cite{ochoa:2015:crossovernetworks}

@@

% Proof sketch
Moves can have their dependencies fixed in some permutation $O$ of all possible moves,
such that $O_0$ is the first move to have its dependencies fixed.
$D_O(i, m)$ is a function which returns how many unfixed dependencies move $m$ has after all
$i$ moves in $O$ before $i$ have been fixed. A greedy solution to this problem works such
that $O_i$ is set to be the move which minimizes $D_O(i, m)$.

All non-greedy solutions must have some $i$ such that $D_O(i, O_i) > D_O(i, O_{i+1})$.
This is because $D_O(i, m)$ can only decrease as $i$ increases, meaning that if some $m$
has a lower value at $i$ than $O_i$, some index between $i$ and when $m$ appears in $O$
this property must hold. Finally, all non-greedy solutions must have some $\hat{i}$
which is the maximum $i$ value for which $D_O(i, O_i) > D_O(i, O_{i+1})$ is true.

Consider the optimal ordering $O^\star$ which is not greedy.
These criteria mean that $D_{O^\star}(\hat{i}, O_{\hat{i}}) > D_{O^\star}(\hat{i}, O_{\hat{i}+1})$.
Swapping $O_{\hat{i}}$ and $O_{\hat{i}+1}$ cannot change the minimum dependency of any other moves.
Performing the swap will improve the minimum dependency of $O_{\hat{i}+1}$ more than it hurts
$O_{\hat{i}}$. Therefore, performing this swap will make $O^\star$ even better, implying it was
not optimal before. Therefore there cannot exist an optimal ordering which is not greedy.

\section{Complexity Classes for Simple Landscapes}

\section{Timing Comparisons}
\subsection{Increasing problem size}
Consider including boxplots of largest size all are successful.

\subsection{The Effect of Radius}
